### 2022-06-10 manuscript.tex
Merge pull request #228 from SORTEE-Github-Hackathon/phpb-filtering-section-a [ci skip]
This build is based on
https://github.com/SORTEE-Github-Hackathon/manuscript/commit/ea3d52568405d25c555daffaef4bbf5a40954c92.

This commit was created by the following CI build and job:
https://github.com/SORTEE-Github-Hackathon/manuscript/commit/ea3d52568405d25c555daffaef4bbf5a40954c92/checks
https://github.com/SORTEE-Github-Hackathon/manuscript/actions/runs/2477668636

This manuscript
(\href{https://SORTEE-Github-Hackathon.github.io/manuscript/v/024357f04df193fe87662943d6e0029321eca2a5/}{permalink})
was automatically generated
from \href{https://github.com/SORTEE-Github-Hackathon/manuscript/tree/024357f04df193fe87662943d6e0029321eca2a5}{SORTEE-Github-Hackathon/manuscript@024357f}
on June 9, 2022.

@@

\hypertarget{practical-ways-of-using-github-in-eeb}{%
\subsection{12 Practical ways of using GitHub in EEB}\label{practical-ways-of-using-github-in-eeb}}

\hypertarget{storing-a-research-compendium}{%
\subsubsection{Storing a research compendium}\label{storing-a-research-compendium}}

In EEB, a research compendium includes all computational materials related to research production, including data, code for analyses and protocols.
Having multiple copies of these files is important to ensure that research can continue should one copy of a file be accidentally modified or deleted.
Many researchers begin using GitHub as a means to backup their research compendium\textsuperscript{\protect\hyperlink{ref-MwwMapRG}{18}} to a remote server (pull-commit-push, see \protect\hyperlink{definitions}{Box 1}).
The GitHub repository serves as a centralized backup and also allows users to synchronize files and work on any device with internet access.
Because GitHub is built on git, version control features are also available, allowing users to re-visit or even restore previous versions of the repository.
Storing a research compendium as a GitHub repository has additional benefits such as facilitating collaboration, integration with data and code archiving services, and contribution to open science, all discussed in sections below.

Unfortunately, GitHub repositories have file size limits, so they may not be appropriate for synchronizing large data files.
Commits (see \protect\hyperlink{definitions}{Box 1}{]}) greater than 50 MB receive a warning and commits larger than 100 MB are blocked (\textsuperscript{\protect\hyperlink{ref-1Co6ZZjF1}{19}}).
Therefore, it is recommended that large datasets are stored only locally or using other cloud services.

\hypertarget{virtual-lab-notebook}{%
\subsubsection{Virtual lab notebook}\label{virtual-lab-notebook}}

Lab notebooks are the virtual or physical notebooks used to help researchers at any career stage keep track of their research methods or laboratory policies\textsuperscript{\protect\hyperlink{ref-wwHxTOtm}{20}}.
Digital lab notebooks stored in the cloud provide clear benefits given the ease with which documents can be shared with new employees and updated as policy changes or experimental methods are modified\textsuperscript{\protect\hyperlink{ref-10V7x4H4l}{21}}.
Increasingly, researchers are leveraging GitHub's underlying version control system to maintain and share digital lab notebooks\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3}}.
At least for aspects of a research project that involve writing code, a GitHub repository is a form of a lab notebook; when changes are made to files in a version controlled repository, the author of those changes makes a commit (\protect\hyperlink{definitions}{Box 1}) accompanied by a message describing the changes and the reason for them. Later, the entire history of commits and their messages are viewable and can be audited similar to a physical lab notebook\textsuperscript{\protect\hyperlink{ref-4ny1onB0}{8}}.
GitHub issues (\protect\hyperlink{definitions}{Box 1}) can be used to prioritize lab objectives and goals, as well as track any status updates.
Some EEB labs have even turned their lab notebooks into shareable websites (\url{https://scheuerell-lab.github.io/lab-book}; \url{https://github.com/HuckleyLab/how_we_work}) as a centralized location for all lab resources.

@@

Modern EEB research is highly collaborative, bringing together multidisciplinary teams from various institutions.
On GitHub, collaborators can share feedback, brainstorm ideas, and troubleshoot problems (Figure \ref{fig:github-diagram}).
GitHub has an ``Issues'' feature that allows for discrete tasks and sub-tasks to be identified, assigned to team members, and categorized with custom labels.
The new GitHub ``Discussion'' feature serves as a message board for conversation.
Scripts, commit messages, and pull requests can be linked directly to issues and discussions, providing a clear record of project workflow.
The use of GitHub for all project-related conversation and planning, rather than e-mail or messaging tools, makes it easier to keep track of progress throughout the lifespan of a project.
Unlike emails and messages which can get lost as more new tasks arise, GitHub issues are intentionally closed by repository administrators hiding the issue from view (closed issues remain accessible but not immediately visible).
Fortunately, it is not essential for all team members to have proficiency in git or programming, as users can interact with Issues and Discussions via web browser or e-mail (e-mail responses still get tracked as comments on the focal GitHub issue).
For larger projects with many team members and tens or hundreds of GitHub issues to sort through, project management software like ZenHub, can help prioritize issues and pull requests.
ZenHub's web interface includes a GitHub Issue visualizer where users can organize issues into high priority or backlogged tasks and link issues together when they are related to a shared project goal or milestone.
GitHub is currently beta testing a similar project management feature called GitHub Projects\textsuperscript{\protect\hyperlink{ref-RhBKe0MG}{22}}.
GitHub can also be integrated with other project management software such as Slack (https://slack.github.com/) so that teams are notified through a Slack group chat when updates are made to a repository.

@@

GitHub supports a broad set of mechanisms for hosting educational materials.
The entire process of running a course, workshop, or lecture, can all be done openly on GitHub including material development, web hosting, and delivery, and even the submission and grading of assignments.
While there are other purpose-built platforms for this, GitHub provides a free, open-source alternative.

Making presentations can be done through most major high-level programming languages such as R, with \texttt{RMarkdown}\textsuperscript{\protect\hyperlink{ref-QqMezOMg}{23}}, Python, with \texttt{python-ppt} (\url{https://python-pptx.readthedocs.io}), and Julia, with \texttt{Remark.jl} (\url{https://juliapackages.com/p/remark}).
Since all these programs work via code bases, they can be version-controlled through git and GitHub.
Once content is made, hosting a course website can be done through GitHub Pages, and there are lots of templates available (e.g., see \url{https://github.com/topics/course-website}).
This way, the course content can be available to enrolled students, as well as a global pool of learners and teachers interested in the course material.
Content can then be delivered via the course website, and/or a GitHub organization with, for example, template repositories for assignments.
Student submissions are perhaps the most challenging component, but the new GitHub Classroom tool (https://classroom.github.com) allows instructors to host private assignments submitted as code files (.R or .Rmd) or .pdf files, and even build custom autograding tests.

Instructors may only need to incorporate some of these tools given the time required to implement various GitHub features, however it is still valuable to do so and to encourage students to begin learning about version control through interacting with git/GitHub, however minimally, through the course.
Especially if a central tenet of a given course or educational unit is to introduce or give students experience with version control, reproducibility, and the tools that working professionals in EEB use, then adopting a few of these tools can be a great way to do so.

@@

Personal or lab websites are one method of improving the dissemination of research findings and coordination of research efforts, but most EEB researchers have little experience in building or hosting webpages.
GitHub Pages (\url{https://pages.github.com}) is a feature that, when activated for a repository, renders content written in markdown (e.g., \url{https://github.com/SORTEE-Github-Hackathon/main-website}) as an HTML web page with a URL (e.g., \url{https://sortee-github-hackathon.github.io/main-website/}).
This can be useful for sharing reports (e.g.~written with RMarkdown and rendered to HTML) with collaborators, or for more complex projects like personal, project, or lab websites.
GitHub also offers website templates and it is possible to create a website by forking (\protect\hyperlink{definitions}{Box 1}) a repository hosting a GitHub Pages site as a starting point.
Aside from free hosting services, another benefit is that GitHub Pages are autogenerated, meaning that when content is modified in the associated GitHub repository, the website instantly updates\textsuperscript{\protect\hyperlink{ref-RVetqmsg}{7}}.
Though the templates are useful for quickly starting up a new website, users are able to fully customize their Pages websites (for technical details of customizing GitHub Pages site see\textsuperscript{\protect\hyperlink{ref-MXxgZJ45}{24}}).
We emphasize that despite the many benefits of using GitHub pages (free hosting, templates, customization), this avenue for creating a website will often be more time intensive than the out-of-the-box platforms. ``Static websites'' (i.e., websites where every user sees the same content) can be created using Jekyll (\url{https://jekyllrb.com}) and Hugo (\url{https://gohugo.io}), which also include template libraries for websites that can be hosted freely via GitHub pages.
Several R packages such as distill (\url{https://github.com/rstudio/distill}) and blogdown (\url{https://bookdown.org/yihui/blogdown}) have made the interface to these website generators more accessible.
In addition, Quarto (\url{https://quarto.org}) has options for creating websites (\url{https://quarto.org/docs/websites}) as well as the code-free website generator wowchemy (\url{https://wowchemy.com}) which uses Hugo.

@@

GitHub makes it easy to store and share a variety of data files in the cloud.
If a repository is made ``public'' the URL to the repository can be shared freely with others.
However, for a variety of reasons (e.g., privately owned company, ability to make repositories private, accounts can be deleted at will) GitHub is not considered a long-term data or code repository like Zenodo and Figshare\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3},\protect\hyperlink{ref-kEX5dgzK}{9},\protect\hyperlink{ref-tbl:compare}{\textbf{tbl:compare?}}}.
Also, unlike long-term repositories, GitHub does not issue Digital Object Identifiers (DOIs) for content uploaded to their servers.
DOIs are persistent and unique alpha-numeric IDs assigned to research products like papers, code, and data.
DOIs allows tracking and citing research products.
For this reason, scientists who share code and data through GitHub are strongly encouraged to also submit GitHub repository content to a long-term data archive\textsuperscript{\protect\hyperlink{ref-1Du6fzB8g}{25}}.
Fortunately, both long-term repositories mentioned above (Zenodo and Figshare) have integrations with GitHub.
After linking one's GitHub repository to Zenodo or Figshare, every time a release (\protect\hyperlink{definitions}{Box 1}) is made, a snapshot of the entire repository is archived with a versioned, citable DOI.
Linking one's GitHub repository with Zenodo, etc. to obtain a DOI helps work become findable, gives proper attribution, and can ensure long-term stability\textsuperscript{\protect\hyperlink{ref-iIEKCTLU}{26}}.

Many researchers believe that their code is not useful because their analysis is context-specific and not explicitly designed for re-use like software.
However, there are many reasons to share data and code beyond re-use.
Even if code lacks detailed annotation, it shows the exact steps taken to conduct an analysis, and therefore provides the most detailed look into how to reproduce a given analysis\textsuperscript{\protect\hyperlink{ref-uBJwnPbq}{27}}.
This is important in light of the reproducibility crisis\textsuperscript{\protect\hyperlink{ref-1HZdsK5Kn}{28}} and will become increasingly important to the collective scientific enterprise as advances in computing power and accessibility unlock the ability to conduct `big data' meta research with data that has already been collected by others.
Failing to include data and code with publications leaves future scientists with fewer open resources from which to understand the world.

@@

GitHub facilitates collaborative coding and asynchronous working because researchers can seamlessly access and contribute to data and code regardless of disparities in schedules or location.
Project work can also be clearly split among team members, giving them the flexibility to contribute when it best fits their schedule.
We acknowledge that there are important differences between the average software developer and EEB researcher using GitHub, and that not all of GitHub's features are useful for research workflows (\protect\hyperlink{tbl:roles}{Table 1}).
However, git's core features such as forking and branching (\protect\hyperlink{definitions}{Box 1}) allow for simultaneous coding on different versions of the same research project, and alternative versions can be discussed and resolved with GitHub.
When EEB projects are contained on the GitHub platform, researchers can stay abreast of progress made by other collaborators without the need for meetings or emails.
Version control\textsuperscript{\protect\hyperlink{ref-K7nbP1Ty}{29}} features also allow users to make progress and changes without worrying about irreparably writing over someone else's work.
By enabling more comprehensive remote collaboration, GitHub encourages the exchange of ideas among researchers at different institutions and in different countries, which can serve to improve the quality of the research itself by providing open access to data and code.
In academic settings, GitHub can also facilitate interactions between research advisors and advisees, providing a platform for students or other trainees to share in-progress code, and flag specific challenges or questions for their supervisors or mentors.
Periodic code review\textsuperscript{\protect\hyperlink{ref-hm9PaCLD}{30}} can also help advisors identify errors early in the process, and inform further training and mentorship to fill gaps in skills.

@@

We wrote this manuscript using Manubot\textsuperscript{\protect\hyperlink{ref-YuJbg3zO}{31}}, a collaborative manuscript platform that uses Markdown for writing and GitHub for storing and tracking changes to a manuscript over time.
Manubot uses a GitHub Actions-based typesetting system to compile individual Markdown files stored in a GitHub repository into a single LaTeX document, which can be displayed in Word, HTML, or PDF formats.
The resulting manuscript can also be compiled using a journal's .tex template to match their formatting requirements.
Since this tool reruns the entire manuscript compilation process with any change to the underlying repository, it can also accomodate continuous integration of code updates into figures and tables with additional GitHub Actions (e.g.~https://github.com/SORTEE-Github-Hackathon/manuscript/tree/main/.github/workflows).
Manubot also allows for straightforward citation management based on URLs or DOIs.
For more examples of Manubot being used for manuscripts, see\textsuperscript{\protect\hyperlink{ref-YuJbg3zO}{31}}.

Since Manubot works on documents in a distributed format, it can be difficult to edit manuscripts for overall flow with only this tool.
We employed hypothes.is (\url{https://hypothes.is}) to write comments on the HTML manuscript document produced by Manubot, which we then addressed by committing changes to the underlying Markdown files via pull requests.
Other authors can reply to the comments to indicate agreement or disagreement, and to note when changes have been made.
Other tools can also be used for version control of scientific manuscripts including R Markdown via the bookdown package (\url{https://bookdown.org/}), jupyter notebooks\textsuperscript{\protect\hyperlink{ref-i4FOZpjN}{32}} and a relatively new tool, Quarto (\url{https://quarto.org}).

@@

Peer review plays a critical role in the scientific evaluation of EEB research.
GitHub provides an open and transparent platform that can be used for either directly providing feedback on research products or building a to-do list from reviewer comments.
Peer review of research code by the open software development community rOpenSci (\url{https://ropensci.org/software-review/}) and of research software and associated manuscripts by the Journal of Open Source Software (\url{https://joss.readthedocs.io/en/latest/submitting.html}) requires that submitted work is hosted on GitHub and their review processes make use of GitHub issues.
rOpenSci's efforts have resulted in many well-used R packages for ecology research including rfishbase\textsuperscript{\protect\hyperlink{ref-Xsdcv6q}{33}} and taxize\textsuperscript{\protect\hyperlink{ref-FVBWKkZu}{34}}.

@@

Conventional research practices typically rely on one or two people running and checking the data analyses, while most coauthors (and readers of the subsequent publication) see only the final results and a verbal description of the analytical steps.
In the developmental stages, collaborators can directly see the code for the analysis, manipulate and explore the data themselves, and check for errors.
Cynically, there is also more insurance against nefarious colleagues that may be tempted to distort results\textsuperscript{\protect\hyperlink{ref-18PTmKJkq}{35}}.
Collaborators as well as pre-publication and post-publication reviewers are better positioned to discover questionable findings if they have full and transparent access to the project.

This transparency can similarly be extended beyond coauthors to the entire scientific community.
Publishing the data and reproducible workflows along with the manuscript allows any reader to review the analysis and reproduce the experiment\textsuperscript{\protect\hyperlink{ref-NOgBWVAr}{36}}.
Supplying code for (novel) methods that are proposed or used also reduces barriers to knowledge and can greatly improve the ability of others to build on existing work, resulting in greater proliferation and accessibility for a broader audience.
GitHub even provides a useful Discussions Forum (\url{https://docs.github.com/en/discussions}) that aids the direct communication with repository owners, as well as the GitHub Community (\url{https://github.community/}) forum for more general questions and sharing of expertise (\protect\hyperlink{tbl:roles}{Table 1}.

@@

The development of research code continues, and so does the need to consider project continuity, especially in EEB where graduate students, research assistants, and postdoctoral fellows often hold relatively short-term positions\textsuperscript{\protect\hyperlink{ref-D4C4k4ak}{37}}.
Often with these types of short-term research projects, once the contract expires, the research code upkeep tends to fall off as the researchers move on to new projects.
Additionally, if the code is kept on only the researcher's personal devices, it becomes increasingly difficult to access the code for future uses.
As projects or contracts end, there should be a handover period of this code in order for the next cohort of researchers to reuse what was already developed\textsuperscript{\protect\hyperlink{ref-4ny1onB0}{8},\protect\hyperlink{ref-D4C4k4ak}{37}}.

GitHub facilitates project continuity among research code by providing tools that make this handover period easier.
As we have already mentioned, using Git for code in Ecology and Evolution can allow for a ``paper trail'' of sorts to be created for the code, thus allowing for future users of the code access to the entire history of the project\textsuperscript{\protect\hyperlink{ref-3DKwn1sY}{14}}.
Additionally, GitHub allows for repositories and organizations to have designated Code Owners\textsuperscript{\protect\hyperlink{ref-s91uGRZ2}{38}}; these code owners can change through time allowing for the transition of code from one cohort of researchers to the next\textsuperscript{\protect\hyperlink{ref-iIEKCTLU}{26}}.
There are also multiple means by which to archive code dependency information with GitHub.
Dependencies can be described in README files, added to environment.yml and/or an environment.txt file, or stored in a repository using tools such as renv (\url{https://rstudio.github.io/renv/articles/renv.html}).
This ensures that as packages continue to develop with new syntax, code will still run as it did when it was developed.

@@

A third barrier may come from general reluctance to share data and code publicly, or technical and logistical issues.
The default public nature of GitHub usage can add additional pressure to students and scientists learning to use the platform.
We also note that large file storage is discouraged (and limited) on GitHub but add-ons do exist (e.g.~\url{https://git-lfs.github.com/}) that permit data storage, and increasing integration between platforms (e.g.\textsuperscript{\protect\hyperlink{ref-lx49NGto}{40}}) allows data to be stored away from GitHub and linked dynamically.
We suspect a major additional barrier to EEB researchers is a distinct lack of GitHub help documents for non-English researchers in ecology and evolution, meaning that EEB researchers potentially miss the opportunity to fully understand the importance of version control, reproducibility, and other benefits of GitHub.

>>>>

This manuscript
(\href{https://SORTEE-Github-Hackathon.github.io/manuscript/v/ea3d52568405d25c555daffaef4bbf5a40954c92/}{permalink})
was automatically generated
from \href{https://github.com/SORTEE-Github-Hackathon/manuscript/tree/ea3d52568405d25c555daffaef4bbf5a40954c92}{SORTEE-Github-Hackathon/manuscript@ea3d525}
on June 10, 2022.

@@

\hypertarget{twelve-practical-ways-github-can-accelerate-research-in-ecology-and-evolution}{%
\subsection{Twelve practical ways GitHub can accelerate research in ecology and evolution}\label{twelve-practical-ways-github-can-accelerate-research-in-ecology-and-evolution}}

\hypertarget{storing-and-sharing-research-compendia}{%
\subsubsection{Storing and sharing research compendia}\label{storing-and-sharing-research-compendia}}

An EEB research compendium includes all computational materials related to research production, including data, code for analyses and protocols.
Having copies of these files safely stored is essential to protect against accidental modifications or deletions.
Many researchers begin using GitHub as a means to store (or backup) their research compendium\textsuperscript{\protect\hyperlink{ref-MwwMapRG}{18}} to a centralized, readily-available remote server (see \protect\hyperlink{definitions}{Box 1}).
A centralized research compendium stored in a version-controlled repository has the advantages of facilitating collaboration, integrating data and code archiving services, allowing file versions to be accessed and restored, and contributes to open science (see sections below).

GitHub limits committed file sizes to 100 Mb (megabytes) (\textsuperscript{\protect\hyperlink{ref-1Co6ZZjF1}{19}}), which can make it challenging for centralizing research compedia containing larger file sizes.
Users may still version large files using Git Large File Storage (LFS;\textsuperscript{\protect\hyperlink{ref-11GtZ7icJ}{20}}/) text pointers, but may have to rely on external file storage alternatives (such as local or cloud-hosting).

\hypertarget{virtual-laboratory-notebooks}{%
\subsubsection{Virtual laboratory notebooks}\label{virtual-laboratory-notebooks}}

Laboratory notebooks help researchers track their research notes, methods, policies and protocols\textsuperscript{\protect\hyperlink{ref-wwHxTOtm}{21}}.
Virtual laboratory notebooks can be stored in GitHub repositories and provide the benefits of simultaneous, centralized and selective access, and allows for the easy update of policies and experiment protocols\textsuperscript{\protect\hyperlink{ref-10V7x4H4l}{22}}.
Researchers have been increasingly using GitHub to maintain versions and share digital laboratory notebooks\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3}} (\emph{e.g.}, \url{https://scheuerell-lab.github.io/lab-book}; \url{https://github.com/HuckleyLab/how_we_work}).

At least for aspects of a research project that involve writing code, a GitHub repository is a form of a laboratory notebook; when changes are made to files in a version controlled repository, the author of those changes makes a commit (\protect\hyperlink{definitions}{Box 1}) accompanied by a description of changes.
Later, the entire history of commits and their commit messages are viewable and can be audited similar to a physical laboratory notebook\textsuperscript{\protect\hyperlink{ref-4ny1onB0}{8}}.
GitHub issues (\protect\hyperlink{definitions}{Box 1}) can be used to prioritize laboratory objectives and goals, as well as track any status updates.

@@

Modern research in ecology and evolution is highly collaborative, bringing together multidisciplinary teams from various institutions.
On GitHub, collaborators can share feedback, brainstorm ideas, and troubleshoot problems (Figure \ref{fig:github-diagram}).
Project management can happen via three GitHub repository features: ``Issues'', ``Discussion'' and ``Projects''.
Github Issues allow for discrete tasks and sub-tasks to be identified, assigned to team members, and categorized with custom labels.
Github Discussions serve as a message board for conversation.
Finally, GitHub Projects integrate issues and pull requests on automated spreadsheets and project boards, providing users with real-time tracking of project priorities and status\textsuperscript{\protect\hyperlink{ref-RhBKe0MG}{23}}.
Scripts, commit messages, and pull requests can be linked directly to issues, discussions, and projects providing a clear record of project workflow.
The use of GitHub for all project-related conversation and planning, rather than e-mail or messaging tools, makes it easier to keep track of progress throughout the lifespan of a project.
Unlike emails and messages which can get lost as more new tasks arise, GitHub issues are intentionally closed by repository administrators hiding the issue from view (closed issues remain accessible but not immediately visible).
Project management in GitHub can also be integrated with third-party applications, such as ZenHub (https://www.zenhub.com/) or Slack (https://slack.github.com/).
ZenHub allows for the enhanced visualization and organization of repositories and their issues, while the GitHub for Slack integration allows notifications from GitHub events to be sent directly to users or group channels\textsuperscript{\protect\hyperlink{ref-1GRWGlDWy}{24}}.

@@

GitHub supports a broad set of mechanisms for hosting educational materials.
The entire process of running a course, workshop, or lecture, can all be done openly on GitHub including material development, web hosting, and delivery, and even submission and grading of assignments.
While there are other purpose-built platforms for this, GitHub provides a free, open-source alternative.

Making presentations, syllabi and other course materials can be done through most major high-level programming languages such as R, with \texttt{RMarkdown}\textsuperscript{\protect\hyperlink{ref-QqMezOMg}{25}}, Python, with \texttt{python-ppt} (\url{https://python-pptx.readthedocs.io}), and Julia, with \texttt{Remark.jl} (\url{https://juliapackages.com/p/remark}), and be version-controlled and stored in GitHub.
Once content is made, hosting a course website can be done through GitHub Pages\textsuperscript{\protect\hyperlink{ref-MDQfMwCW}{26}} (\emph{e.g.}, https://github.com/topics/course-website\textgreater).
This way, the course content can be available to enrolled students, as well as a global pool of learners and teachers interested in the course material.
Content can then be delivered via the course website, and/or a GitHub Organization with, for example, template repositories for assignments.
Student submissions are perhaps the most challenging component, but the new GitHub Classroom tool (\url{https://classroom.github.com}) allows instructors to host private assignments to be submitted collaboratively or individually as code or PDF files, and even build autograding tests.
Although time-consuming to establish, using these features can integrate learning version control and GitHub with the learning course content, and thus boost students feelings of self-efficacy and confidence\textsuperscript{\protect\hyperlink{ref-dqrFjoSb}{27}}.

@@

Personal or lab websites are one method of improving the dissemination of research findings and coordination of research efforts, but most EEB researchers have little experience in building or hosting webpages.
GitHub Pages (\url{https://pages.github.com}) is a feature that, when activated for a repository, renders content written in markdown (e.g., \url{https://github.com/SORTEE-Github-Hackathon/main-website}) as an HTML web page with a URL (e.g., \url{https://sortee-github-hackathon.github.io/main-website/}).
This can be useful for sharing reports (e.g.~written with RMarkdown and rendered to HTML) with collaborators, or for more complex projects like personal, project, or lab websites.
GitHub also offers website templates and it is possible to create a website by forking (\protect\hyperlink{definitions}{Box 1}) a repository hosting a GitHub Pages site as a starting point.
Aside from free hosting services, another benefit is that GitHub Pages are autogenerated, meaning that when content is modified in the associated GitHub repository, the website instantly updates\textsuperscript{\protect\hyperlink{ref-RVetqmsg}{7}}.
Though the templates are useful for quickly starting up a new website, users are able to fully customize their Pages websites (for technical details of customizing GitHub Pages site see\textsuperscript{\protect\hyperlink{ref-MXxgZJ45}{28}}).
We emphasize that despite the many benefits of using GitHub pages (free hosting, templates, customization), this avenue for creating a website will often be more time intensive than the out-of-the-box platforms. ``Static websites'' (i.e., websites where every user sees the same content) can be created using Jekyll (\url{https://jekyllrb.com}) and Hugo (\url{https://gohugo.io}), which also include template libraries for websites that can be hosted freely via GitHub pages.
Several R packages such as distill (\url{https://github.com/rstudio/distill}) and blogdown (\url{https://bookdown.org/yihui/blogdown}) have made the interface to these website generators more accessible.
In addition, Quarto (\url{https://quarto.org}) has options for creating websites (\url{https://quarto.org/docs/websites}) as well as the code-free website generator wowchemy (\url{https://wowchemy.com}) which uses Hugo.

@@

GitHub makes it easy to store and share a variety of data files in the cloud.
If a repository is made ``public'' the URL to the repository can be shared freely with others.
However, for a variety of reasons (e.g., privately owned company, ability to make repositories private, accounts can be deleted at will) GitHub is not considered a long-term data or code repository like Zenodo and Figshare\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3},\protect\hyperlink{ref-kEX5dgzK}{9},\protect\hyperlink{ref-tbl:compare}{\textbf{tbl:compare?}}}.
Also, unlike long-term repositories, GitHub does not issue Digital Object Identifiers (DOIs) for content uploaded to their servers.
DOIs are persistent and unique alpha-numeric IDs assigned to research products like papers, code, and data.
DOIs allows tracking and citing research products.
For this reason, scientists who share code and data through GitHub are strongly encouraged to also submit GitHub repository content to a long-term data archive\textsuperscript{\protect\hyperlink{ref-1Du6fzB8g}{29}}.
Fortunately, both long-term repositories mentioned above (Zenodo and Figshare) have integrations with GitHub.
After linking one's GitHub repository to Zenodo or Figshare, every time a release (\protect\hyperlink{definitions}{Box 1}) is made, a snapshot of the entire repository is archived with a versioned, citable DOI.
Linking one's GitHub repository with Zenodo, etc. to obtain a DOI helps work become findable, gives proper attribution, and can ensure long-term stability\textsuperscript{\protect\hyperlink{ref-iIEKCTLU}{30}}.

Many researchers believe that their code is not useful because their analysis is context-specific and not explicitly designed for re-use like software.
However, there are many reasons to share data and code beyond re-use.
Even if code lacks detailed annotation, it shows the exact steps taken to conduct an analysis, and therefore provides the most detailed look into how to reproduce a given analysis\textsuperscript{\protect\hyperlink{ref-uBJwnPbq}{31}}.
This is important in light of the reproducibility crisis\textsuperscript{\protect\hyperlink{ref-1HZdsK5Kn}{32}} and will become increasingly important to the collective scientific enterprise as advances in computing power and accessibility unlock the ability to conduct `big data' meta research with data that has already been collected by others.
Failing to include data and code with publications leaves future scientists with fewer open resources from which to understand the world.

@@

GitHub facilitates collaborative coding and asynchronous working because researchers can seamlessly access and contribute to data and code regardless of disparities in schedules or location.
Project work can also be clearly split among team members, giving them the flexibility to contribute when it best fits their schedule.
We acknowledge that there are important differences between the average software developer and EEB researcher using GitHub, and that not all of GitHub's features are useful for research workflows (\protect\hyperlink{tbl:roles}{Table 1}).
However, git's core features such as forking and branching (\protect\hyperlink{definitions}{Box 1}) allow for simultaneous coding on different versions of the same research project, and alternative versions can be discussed and resolved with GitHub.
When EEB projects are contained on the GitHub platform, researchers can stay abreast of progress made by other collaborators without the need for meetings or emails.
Version control\textsuperscript{\protect\hyperlink{ref-K7nbP1Ty}{33}} features also allow users to make progress and changes without worrying about irreparably writing over someone else's work.
By enabling more comprehensive remote collaboration, GitHub encourages the exchange of ideas among researchers at different institutions and in different countries, which can serve to improve the quality of the research itself by providing open access to data and code.
In academic settings, GitHub can also facilitate interactions between research advisors and advisees, providing a platform for students or other trainees to share in-progress code, and flag specific challenges or questions for their supervisors or mentors.
Periodic code review\textsuperscript{\protect\hyperlink{ref-hm9PaCLD}{34}} can also help advisors identify errors early in the process, and inform further training and mentorship to fill gaps in skills.

@@

We wrote this manuscript using Manubot\textsuperscript{\protect\hyperlink{ref-YuJbg3zO}{35}}, a collaborative manuscript platform that uses Markdown for writing and GitHub for storing and tracking changes to a manuscript over time.
Manubot uses a GitHub Actions-based typesetting system to compile individual Markdown files stored in a GitHub repository into a single LaTeX document, which can be displayed in Word, HTML, or PDF formats.
The resulting manuscript can also be compiled using a journal's .tex template to match their formatting requirements.
Since this tool reruns the entire manuscript compilation process with any change to the underlying repository, it can also accomodate continuous integration of code updates into figures and tables with additional GitHub Actions (e.g.~https://github.com/SORTEE-Github-Hackathon/manuscript/tree/main/.github/workflows).
Manubot also allows for straightforward citation management based on URLs or DOIs.
For more examples of Manubot being used for manuscripts, see\textsuperscript{\protect\hyperlink{ref-YuJbg3zO}{35}}.

Since Manubot works on documents in a distributed format, it can be difficult to edit manuscripts for overall flow with only this tool.
We employed hypothes.is (\url{https://hypothes.is}) to write comments on the HTML manuscript document produced by Manubot, which we then addressed by committing changes to the underlying Markdown files via pull requests.
Other authors can reply to the comments to indicate agreement or disagreement, and to note when changes have been made.
Other tools can also be used for version control of scientific manuscripts including R Markdown via the bookdown package (\url{https://bookdown.org/}), jupyter notebooks\textsuperscript{\protect\hyperlink{ref-i4FOZpjN}{36}} and a relatively new tool, Quarto (\url{https://quarto.org}).

@@

Peer review plays a critical role in the scientific evaluation of EEB research.
GitHub provides an open and transparent platform that can be used for either directly providing feedback on research products or building a to-do list from reviewer comments.
Peer review of research code by the open software development community rOpenSci (\url{https://ropensci.org/software-review/}) and of research software and associated manuscripts by the Journal of Open Source Software (\url{https://joss.readthedocs.io/en/latest/submitting.html}) requires that submitted work is hosted on GitHub and their review processes make use of GitHub issues.
rOpenSci's efforts have resulted in many well-used R packages for ecology research including rfishbase\textsuperscript{\protect\hyperlink{ref-Xsdcv6q}{37}} and taxize\textsuperscript{\protect\hyperlink{ref-FVBWKkZu}{38}}.

@@

Conventional research practices typically rely on one or two people running and checking the data analyses, while most coauthors (and readers of the subsequent publication) see only the final results and a verbal description of the analytical steps.
In the developmental stages, collaborators can directly see the code for the analysis, manipulate and explore the data themselves, and check for errors.
Cynically, there is also more insurance against nefarious colleagues that may be tempted to distort results\textsuperscript{\protect\hyperlink{ref-18PTmKJkq}{39}}.
Collaborators as well as pre-publication and post-publication reviewers are better positioned to discover questionable findings if they have full and transparent access to the project.

This transparency can similarly be extended beyond coauthors to the entire scientific community.
Publishing the data and reproducible workflows along with the manuscript allows any reader to review the analysis and reproduce the experiment\textsuperscript{\protect\hyperlink{ref-NOgBWVAr}{40}}.
Supplying code for (novel) methods that are proposed or used also reduces barriers to knowledge and can greatly improve the ability of others to build on existing work, resulting in greater proliferation and accessibility for a broader audience.
GitHub even provides a useful Discussions Forum (\url{https://docs.github.com/en/discussions}) that aids the direct communication with repository owners, as well as the GitHub Community (\url{https://github.community/}) forum for more general questions and sharing of expertise (\protect\hyperlink{tbl:roles}{Table 1}.

@@

The development of research code continues, and so does the need to consider project continuity, especially in EEB where graduate students, research assistants, and postdoctoral fellows often hold relatively short-term positions\textsuperscript{\protect\hyperlink{ref-D4C4k4ak}{41}}.
Often with these types of short-term research projects, once the contract expires, the research code upkeep tends to fall off as the researchers move on to new projects.
Additionally, if the code is kept on only the researcher's personal devices, it becomes increasingly difficult to access the code for future uses.
As projects or contracts end, there should be a handover period of this code in order for the next cohort of researchers to reuse what was already developed\textsuperscript{\protect\hyperlink{ref-4ny1onB0}{8},\protect\hyperlink{ref-D4C4k4ak}{41}}.

GitHub facilitates project continuity among research code by providing tools that make this handover period easier.
As we have already mentioned, using Git for code in Ecology and Evolution can allow for a ``paper trail'' of sorts to be created for the code, thus allowing for future users of the code access to the entire history of the project\textsuperscript{\protect\hyperlink{ref-3DKwn1sY}{14}}.
Additionally, GitHub allows for repositories and organizations to have designated Code Owners\textsuperscript{\protect\hyperlink{ref-s91uGRZ2}{42}}; these code owners can change through time allowing for the transition of code from one cohort of researchers to the next\textsuperscript{\protect\hyperlink{ref-iIEKCTLU}{30}}.
There are also multiple means by which to archive code dependency information with GitHub.
Dependencies can be described in README files, added to environment.yml and/or an environment.txt file, or stored in a repository using tools such as renv (\url{https://rstudio.github.io/renv/articles/renv.html}).
This ensures that as packages continue to develop with new syntax, code will still run as it did when it was developed.

@@

A third barrier may come from general reluctance to share data and code publicly, or technical and logistical issues.
The default public nature of GitHub usage can add additional pressure to students and scientists learning to use the platform.
We also note that large file storage is discouraged (and limited) on GitHub but add-ons do exist (e.g.~\url{https://git-lfs.github.com/}) that permit data storage, and increasing integration between platforms (e.g.\textsuperscript{\protect\hyperlink{ref-lx49NGto}{44}}) allows data to be stored away from GitHub and linked dynamically.
We suspect a major additional barrier to EEB researchers is a distinct lack of GitHub help documents for non-English researchers in ecology and evolution, meaning that EEB researchers potentially miss the opportunity to fully understand the importance of version control, reproducibility, and other benefits of GitHub.
